---
title: "Analysis of the phosphoTPP dataset by Potel et al."
author: "Nils Kurzawa"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: BiocStyle::pdf_document
bibliography: references.bib
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Analysis of the phosphoTPP dataset by Potel et al.}
  %VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
reRunFitting <- FALSE
exportPath <- "."
```

# Introduction

This vignette illustrates the analysis of the phosphoTPP dataset by @potel_2020 and compares it to the one by @huang_2019.

# Step-by-step walk through the analysis

## Download of required data sets

The supplementary tables `41592_2019_499_MOESM3_ESM.xlsx` and `41592_2019_499_MOESM4_ESM.xlsx` by @huang_2019 can be downloaded from the journal website: https://www.nature.com/articles/s41592-019-0499-3#Sec26 .

The supplementary tables `56035_2_data_set_478471_pxb8gk.xlsx` and `56035_2_data_set_478475_pxb8l1-tableS3.xlsx` by @ochoa_2020 can also be downloaded from the respective journal website: https://www.nature.com/articles/s41587-019-0344-3#Sec30 .

The processed data by @potel_2020 can be downloaded from Mendeley: LINK TO BE INSERTED

## Setup

Install the `phosphoTPP` package from github
```{r eval=FALSE}
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("nkurzaw/phosphoTPP")
```

Set some plotting options
```{r}
# set plotting sizes
anno_size <- 2
pval_size <- 2
rsq_size <- 2

# define plotting theme
theme_paper <- 
  theme(axis.title = element_text(size = 8),
        axis.text = element_text(size = 6),
        legend.text = element_text(size = 6),
        title = element_text(size = 9))
```

We then load the required libraries
```{r, warning=FALSE, message=FALSE}
library(phosphoTPP)
library(dplyr)
library(tidyr)
library(GGally)
library(ggpmisc)
```

And we set the thresholds for filtering for high-quality peptide identifications and quantifications and the filtering criteria for melting curves.
```{r}
# set thresholds for quality filtering
s2i_thres <-  0.5
p2t_thres <-  4
mascot_score_thres <- 20
fdr_at_score_thres <- 0.01
r2_thres <- 0.8
plateau_thres <- 0.2
```

Load the config table which tells us which TMT channels were used for which temperature treatment
```{r}
cfg_tab <- readxl::read_xlsx(file.path("../dat", "TPP-TR_config_phospho.xlsx"))
cfg_tab
```

```{r, eval=FALSE}
temperature_anno <- cfg_tab %>%
  dplyr::select(matches('[0-9]{3}')) %>%
  gather(key, temperature) %>%
  mutate(channel = paste("sig", key, sep = "_")) %>% 
  mutate(fc_channel = paste("rel_fc", key, sep = "_"))

temperature_anno_prot <- cfg_tab  %>%
  dplyr::select(matches('[0-9]{3}')) %>%
  gather(key, temperature) %>%
  mutate(channel = paste("signal_sum", key, sep = "_")) %>% 
  mutate(fc_channel = paste("rel_fc", key, sep = "_"))
```

```{r, eval=FALSE}
# read in lists of phospho and nbt protein and peptides tables per replicate
phospho_protein_files <- list.files("../dat", pattern = 'phospho.+protein', full.names = TRUE)
phospho_protein_files
```

```{r, eval=FALSE}
nbf_protein_files <- list.files("../dat", pattern = 'NBF.+protein', full.names = TRUE)
nbf_protein_files
```

```{r, eval=FALSE}
phospho_peptide_files <- list.files("../dat", pattern = 'phospho.+peptide.+txt', full.names = TRUE)
phospho_peptide_files
```

```{r, eval=FALSE}
phosho_mxq_peptide_files <- list.files("../dat", pattern = 'pTPP.+txt', full.names = TRUE)
phosho_mxq_peptide_files
```

```{r, eval=FALSE}
nbf_peptide_files <- list.files("../dat", pattern = 'NBF.+peptide', full.names = TRUE)
nbf_peptide_files
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
# read data files into lists of replicates
phospho_protein_tabs <- lapply(seq_len(length(phospho_protein_files)), function(file_i){
  read_delim(phospho_protein_files[file_i], delim = "\t") %>% 
    mutate(replicate = file_i)
})

nbf_protein_tabs <- lapply(seq_len(length(nbf_protein_files)), function(file_i){
  read_delim(nbf_protein_files[file_i], delim = "\t") %>% 
    mutate(replicate = file_i)
})

phospho_peptide_tabs <- lapply(seq_len(length(phospho_peptide_files)), function(file_i){
  ibq_tab <- read_delim(phospho_peptide_files[file_i], delim = "\t") %>% 
    filter_at(vars(starts_with("sig_")), all_vars(!is.na(.))) 
  mxq_tab <- read_delim(phosho_mxq_peptide_files[file_i], delim = "\t")
  
  combined_tab <- left_join(
  ibq_tab %>% dplyr::select(protein_id, sequence, modifications,
                            matches("sig_"), s2i, p2t, msms_id),
  mxq_tab %>% dplyr::select(protein_id_mxq = Proteins, sequence = Sequence, msms_id = `MS/MS scan number`,
                            mod_sequence = `Modified sequence`, phospho_prob = `Phospho (STY) Probabilities`,
                            phospho_score_diff = `Phospho (STY) Score Diffs`, phospho_site_STY = `Phospho (STY)`),
  by = c("sequence", "msms_id")) %>%
  filter(!is.na(protein_id_mxq)) %>%
  group_by(sequence, msms_id) %>%
  filter(!duplicated(mod_sequence)) %>%
  ungroup %>%
  filter(getProb(phospho_prob)) %>%
  filter(p2t >= p2t_thres, s2i >= s2i_thres) %>% 
  mutate(replicate = file_i) %>% 
    mutate(rel_fc_126 = sig_126 / sig_126,
           rel_fc_127L = sig_127L / sig_126,
           rel_fc_127H = sig_127H / sig_126,
           rel_fc_128L = sig_128L / sig_126,
           rel_fc_128H = sig_128H / sig_126,
           rel_fc_129L = sig_129L / sig_126,
           rel_fc_129H = sig_129H / sig_126,
           rel_fc_130L = sig_130L / sig_126,
           rel_fc_130H = sig_130H / sig_126,
           rel_fc_131L = sig_131L / sig_126)
  
  return(combined_tab)
})

phospho_non_phospho_peptide_tabs <- lapply(seq_len(length(phospho_peptide_files)), function(file_i){
  read_delim(phospho_peptide_files[file_i], delim = "\t") %>% 
    filter(is_decoy == 0, score > mascot_score_thres & is_unique == 1, rank == 1,
           fdr_at_score < fdr_at_score_thres, !is.na(modifications), 
           grepl("TMT", modifications), !grepl("Phospho", modifications),
           in_quantification_of_protein == 1, !grepl("##", protein_id)) %>%
    filter_at(vars(starts_with("sig_")), all_vars(!is.na(.))) %>% 
    mutate(replicate = file_i) %>% 
    mutate(rel_fc_126 = sig_126 / sig_126,
           rel_fc_127L = sig_127L / sig_126,
           rel_fc_127H = sig_127H / sig_126,
           rel_fc_128L = sig_128L / sig_126,
           rel_fc_128H = sig_128H / sig_126,
           rel_fc_129L = sig_129L / sig_126,
           rel_fc_129H = sig_129H / sig_126,
           rel_fc_130L = sig_130L / sig_126,
           rel_fc_130H = sig_130H / sig_126,
           rel_fc_131L = sig_131L / sig_126)
})

nbf_peptide_tabs <- lapply(seq_len(length(nbf_peptide_files)), function(file_i){
  read_delim(nbf_peptide_files[file_i], delim = "\t") %>% 
    filter(is_decoy == 0, score > mascot_score_thres & is_unique == 1, rank == 1,
           fdr_at_score < fdr_at_score_thres, !is.na(modifications), 
           grepl("TMT", modifications), 
           in_quantification_of_protein == 1, !grepl("##", protein_id)) %>%
    filter_at(vars(starts_with("sig_")), all_vars(!is.na(.))) %>% 
    mutate(replicate = file_i) %>% 
    mutate(rel_fc_126 = sig_126 / sig_126,
           rel_fc_127L = sig_127L / sig_126,
           rel_fc_127H = sig_127H / sig_126,
           rel_fc_128L = sig_128L / sig_126,
           rel_fc_128H = sig_128H / sig_126,
           rel_fc_129L = sig_129L / sig_126,
           rel_fc_129H = sig_129H / sig_126,
           rel_fc_130L = sig_130L / sig_126,
           rel_fc_130H = sig_130H / sig_126,
           rel_fc_131L = sig_131L / sig_126)
})

```

```{r, eval=FALSE}
# compute normalization factors
## first compute median normalization factors for shared non-phospho peptides between 
## matching phospho-enriched and nbf samples
n_rep <- min(length(phospho_peptide_files),
             length(nbf_peptide_files))

per_exp_norm_factors <- lapply(seq_len(n_rep), function(rep_i){
  norm_in_df <- bind_rows(nbf_peptide_tabs[[rep_i]] %>% 
    mutate(dataset_id = paste("nbf", replicate, sep = "_")),
    phospho_non_phospho_peptide_tabs[[rep_i]] %>% 
    mutate(dataset_id = paste("phospho", replicate, sep = "_")))
  
  nbf_peptide_norm_factors <- getNormFactors4Data(
    pep_tab = norm_in_df, 
    temperature_anno = temperature_anno,
    filter_criteria = list("rel_fc_127H_low" = 0,
                           "rel_fc_129H_low" = 0.4,
                           "rel_fc_129H_high" = 0.6,
                           "rel_fc_130H_low" = 0,
                           "rel_fc_130H_high" = 0.3,
                           "rel_fc_131L_low" = 0,
                           "rel_fc_131L_high" = 0.2),
    n_rep = 2)
  
  median_norm_factors <- nbf_peptide_norm_factors
  out_df <- tibble(
    temperature = nbf_peptide_norm_factors[[1]]$temperature,
    median_norm_factor = nbf_peptide_norm_factors[[1]]$median_rel_value /
      nbf_peptide_norm_factors[[2]]$median_rel_value)
  
  return(out_df)
})

per_exp_norm_factors
```

```{r, eval=FALSE}
## now compute normalizing factors retrieved from curves fitted on hq data shared 
## between different nbf replicates
norm_in_df <- 
  bind_rows(nbf_peptide_tabs) %>% 
    mutate(dataset_id = paste("nbf", replicate, sep = "_"))#,

nbf_peptide_norm_factors <- getNormFactors4Data(
  pep_tab = norm_in_df, 
  temperature_anno = temperature_anno,
      filter_criteria = list("rel_fc_127H_low" = 1,
                           "rel_fc_129H_low" = 0.4,
                           "rel_fc_129H_high" = 0.6,
                           "rel_fc_130H_low" = 0,
                           "rel_fc_130H_high" = 0.3,
                           "rel_fc_131L_low" = 0,
                           "rel_fc_131L_high" = 0.2),
  n_rep = n_rep)

nbf_peptide_norm_factors
```

```{r, eval=FALSE}
# retrieve one table of per normalized nbf/phospho peptide fold changes
# and compare to non-normalized one
nbf_peptide_filtered_not_norm <- bind_rows(lapply(seq_len(length(nbf_peptide_tabs)),
                                       function(tab_i){
  nbf_peptide_tabs[[tab_i]] %>%
    dplyr::select(protein_id, sequence, modifications,
                  matches("rel_fc_"), replicate, score) %>%
    gather(key, value, -protein_id, -sequence, -modifications,
           -replicate, -score) %>%
    group_by(protein_id, sequence, replicate, modifications, key) %>%
    filter(score == max(score)) %>%
    left_join(temperature_anno, by = c("key" = "fc_channel")) %>%
    mutate(rel_value = value * 1) %>% 
    left_join(nbf_protein_tabs[[tab_i]] %>% dplyr::select(protein_id, gene_name),
              by = c("protein_id")) 
}))

nbf_peptide_filtered <- bind_rows(lapply(seq_len(length(nbf_peptide_tabs)),
                                       function(tab_i){
  nbf_peptide_tabs[[tab_i]] %>%
    dplyr::select(protein_id, sequence, modifications,
                  matches("rel_fc_"), replicate, score) %>%
    gather(key, value, -protein_id, -sequence, -modifications,
           -replicate, -score) %>%
    group_by(protein_id, sequence, replicate, modifications, key) %>%
    filter(score == max(score)) %>%
    left_join(temperature_anno, by = c("key" = "fc_channel")) %>%
    left_join(nbf_peptide_norm_factors[[tab_i]],
              by = "temperature") %>% 
    mutate(rel_value = value * norm_factor) %>% 
    left_join(nbf_protein_tabs[[tab_i]] %>% dplyr::select(protein_id, gene_name),
              by = c("protein_id")) 
}))

phospho_peptide_tab_filtered_not_norm <- bind_rows(lapply(seq_len(length(phospho_peptide_tabs)), function(tab_i){
  phospho_peptide_tabs[[tab_i]] %>%
    dplyr::select(-matches("sig_")) %>% 
    gather(key, value, matches("rel_fc_")) %>%
    group_by(protein_id, sequence, mod_sequence, phospho_site_STY, 
             replicate, key) %>%
    summarise(value = mean(value, na.rm = TRUE),
              mean_s2i = mean(s2i, na.rm = TRUE),
              mean_p2t = mean(p2t, na.rm = TRUE)) %>%
    ungroup %>%
    left_join(temperature_anno %>% dplyr::select(-key),
              by = c("key" = "fc_channel")) %>%
    mutate(rel_value = value * 1) %>%
    left_join(phospho_protein_tabs[[tab_i]] %>% 
                dplyr::select(protein_id, gene_name),
              by = c("protein_id")) %>%
    filter(!grepl("##", protein_id))
}))

phospho_peptide_tab_filtered <- bind_rows(lapply(seq_len(length(phospho_peptide_tabs)), function(tab_i){
  phospho_peptide_tabs[[tab_i]] %>%
    dplyr::select(-matches("sig_")) %>% 
    gather(key, value, matches("rel_fc_")) %>%
    group_by(protein_id, sequence, mod_sequence, phospho_site_STY, 
             replicate, key) %>%
    summarise(value = mean(value, na.rm = TRUE),
              mean_s2i = mean(s2i, na.rm = TRUE),
              mean_p2t = mean(p2t, na.rm = TRUE)) %>%
    ungroup %>%
    left_join(temperature_anno %>% dplyr::select(-key),
              by = c("key" = "fc_channel")) %>%
    left_join(per_exp_norm_factors[[tab_i]],
              by = "temperature") %>% 
    left_join(nbf_peptide_norm_factors[[tab_i]],
              by = "temperature") %>% 
    mutate(rel_value = value * median_norm_factor * norm_factor) %>%
    left_join(phospho_protein_tabs[[tab_i]] %>% 
                dplyr::select(protein_id, gene_name),
              by = c("protein_id")) %>%
    filter(!grepl("##", protein_id))
}))

p_qc1_1 <- nbf_peptide_filtered_not_norm %>% 
  ggplot(aes(temperature, rel_value)) + 
  geom_boxplot(aes(group = temperature)) + 
  facet_wrap(~replicate) + 
  coord_cartesian(ylim = c(0, 2.5))

p_qc1_3 <- nbf_peptide_filtered %>% 
  ggplot(aes(temperature, rel_value)) + 
  geom_boxplot(aes(group = temperature)) + 
  facet_wrap(~replicate) + 
  coord_cartesian(ylim = c(0, 2.5))

p_qc1_2 <- phospho_peptide_tab_filtered_not_norm %>% 
  ggplot(aes(temperature, rel_value)) + 
  geom_boxplot(aes(group = temperature)) + 
  facet_wrap(~replicate) + 
  coord_cartesian(ylim = c(0, 2.5))

p_qc1_4 <- phospho_peptide_tab_filtered %>% 
  ggplot(aes(temperature, rel_value)) + 
  geom_boxplot(aes(group = temperature)) + 
  facet_wrap(~replicate) + 
  coord_cartesian(ylim = c(0, 2.5))

cowplot::plot_grid(p_qc1_1 + ggtitle("raw nbf"), p_qc1_2 + ggtitle("raw phospho"), 
          p_qc1_3 + ggtitle("curve-based norm nbf"),
          p_qc1_4 + ggtitle("curve-based norm phospho, whith prior median norm by non-phospho peptides"), ncol = 2)
```

```{r, warning=FALSE, message=FALSE, eval=reRunFitting}
# fit curves to nbf dataset
out_nbf_df <- nbf_peptide_filtered %>% 
  group_by(gene_name, replicate) %>%
  do({
    suppressWarnings(fitMeltcurveModelAndEval(df = .))
  }) %>% 
  ungroup
```

```{r eval=!reRunFitting, include=FALSE}
out_nbf_df <- readRDS("../prerun/out_nbf_df.rds")
```

```{r}
# filter for high quality fits
nbf_tm_rep_df <- out_nbf_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  dplyr::select(gene_name, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)

pm <- GGally::ggpairs(nbf_tm_rep_df, 2:6,
        upper = "blank", diag = NULL,
        lower = list(continuous = GGally::wrap(lowerFn)))

GGally::ggmatrix(list(pm[2, 1], NULL, NULL, NULL,
              pm[3, 1], pm[3,2], NULL, NULL,
              pm[4, 1], pm[4,2], pm[4,3], NULL,
              pm[5, 1], pm[5,2], pm[5,3], pm[5,4]), 4, 4,
         xAxisLabels = c("Tm rep1", "Tm rep2", "Tm rep3", "Tm rep4"),
         yAxisLabels = c("Tm rep2", "Tm rep3","Tm rep4","Tm rep5"))
```


```{r, warning=FALSE, message=FALSE, eval=reRunFitting}
# fit curves to not-normalized nbf dataset
out_nbf_nn_df <- nbf_peptide_filtered_not_norm %>% 
  group_by(gene_name, replicate) %>%
  do({
    suppressWarnings(fitMeltcurveModelAndEval(df = .))
  }) %>% 
  ungroup
```

```{r eval=!reRunFitting, include=FALSE}
out_nbf_nn_df <- readRDS("../prerun/out_nbf_nn_df.rds")
```

```{r}
# filter for high quality fits
nbf_tm_rep_nn_df <- out_nbf_nn_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  dplyr::select(gene_name, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)


pm_nn <- ggpairs(nbf_tm_rep_nn_df, 2:6,
        upper = "blank", diag = NULL,
        lower = list(continuous = wrap(lowerFn)))

ggmatrix(list(pm_nn[2, 1], NULL, NULL, NULL,
              pm_nn[3, 1], pm_nn[3,2], NULL, NULL,
              pm_nn[4, 1], pm_nn[4,2], pm_nn[4,3], NULL,
              pm_nn[5, 1], pm_nn[5,2], pm_nn[5,3], pm_nn[5,4]), 4, 4,
         xAxisLabels = c("Tm rep1", "Tm rep2", "Tm rep3", "Tm rep4"),
         yAxisLabels = c("Tm rep2", "Tm rep3","Tm rep4","Tm rep5"))
```

```{r, warning=FALSE, message=FALSE, eval=reRunFitting}
# fit phospho peptide melting curves
out_phospho_df <- phospho_peptide_tab_filtered %>%
  dplyr::select(protein_id, gene_name, sequence, 
                mod_sequence, mean_s2i, mean_p2t,
                phospho_site_STY, replicate, 
                rel_value, temperature) %>%
  group_by(protein_id, gene_name, sequence, 
           mod_sequence, mean_s2i, mean_p2t,
           phospho_site_STY, replicate) %>%
  do({
    suppressWarnings(fitMeltcurveModelAndEval(df = .))
  }) %>% 
  ungroup
```

```{r eval=!reRunFitting, include=FALSE}
out_phospho_df <- readRDS("../prerun/out_phospho_df.rds")
```


```{r}
phospho_tm_rep_df <- out_phospho_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  dplyr::select(gene_name, sequence, mod_sequence, phospho_site_STY, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)

pm_phospho <- ggpairs(phospho_tm_rep_df, 5:9,
        upper = "blank", diag = NULL,
        lower = list(continuous = wrap(lowerFn)))

ggmatrix(list(pm_phospho[2, 1], NULL, NULL, NULL,
              pm_phospho[3, 1], pm_phospho[3,2], NULL, NULL,
              pm_phospho[4, 1], pm_phospho[4,2], pm_phospho[4,3], NULL,
              pm_phospho[5, 1], pm_phospho[5,2], pm_phospho[5,3], pm_phospho[5,4]), 4, 4,
         xAxisLabels = c("Tm rep1", "Tm rep2", "Tm rep3", "Tm rep4"),
         yAxisLabels = c("Tm rep2", "Tm rep3","Tm rep4","Tm rep5"))
```

```{r, warning=FALSE, message=FALSE, eval=FALSE}
# fit non-norm phospho peptide melting curves
out_phospho_nn_df <- phospho_peptide_tab_filtered_not_norm %>%
  dplyr::select(protein_id, gene_name, sequence, 
                mod_sequence, mean_s2i, mean_p2t,
                phospho_site_STY, replicate, 
                rel_value, temperature) %>%
  group_by(protein_id, gene_name, sequence, 
           mod_sequence, mean_s2i, mean_p2t,
           phospho_site_STY, replicate) %>%
  do({
    suppressWarnings(fitMeltcurveModelAndEval(df = .))
  }) %>% 
  ungroup
```

```{r eval=!reRunFitting, include=FALSE}
out_phospho_nn_df <- readRDS("../prerun/out_phospho_nn_df.rds")
```


```{r}
phospho_tm_rep_nn_df <- out_phospho_nn_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  dplyr::select(gene_name, sequence, mod_sequence, 
                phospho_site_STY, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)

pm_phospho_nn <- GGally::ggpairs(phospho_tm_rep_nn_df, 5:9,
        upper = "blank", diag = NULL,
        lower = list(continuous = wrap(lowerFn)))

GGally::ggmatrix(list(pm_phospho_nn[2, 1], NULL, NULL, NULL,
              pm_phospho_nn[3, 1], pm_phospho_nn[3,2], NULL, NULL,
              pm_phospho_nn[4, 1], pm_phospho_nn[4,2], pm_phospho_nn[4,3], NULL,
              pm_phospho_nn[5, 1], pm_phospho_nn[5,2], pm_phospho_nn[5,3], pm_phospho_nn[5,4]), 4, 4,
         xAxisLabels = c("Tm rep1", "Tm rep2", "Tm rep3", "Tm rep4"),
         yAxisLabels = c("Tm rep2", "Tm rep3","Tm rep4","Tm rep5"))
```

```{r, eval=FALSE}
# create scatterplot of phospho and nbf Tm with non-normalized data
phospho_combo_nn_df <- left_join(
  out_phospho_nn_df %>%
    filter(rSq > r2_thres, plateau < plateau_thres) %>% 
    group_by(protein_id, gene_name, sequence, 
             mod_sequence, phospho_site_STY) %>% 
    summarise(meltPoint = mean(meltPoint, na.rm = TRUE),
              plateau = mean(plateau, na.rm = TRUE),
              rSq = mean(rSq, na.rm = TRUE),
              n_rep = n()) %>%
    ungroup,
  out_nbf_nn_df %>% 
    group_by(gene_name) %>% 
    filter(rSq > r2_thres, plateau < plateau_thres) %>% 
    summarise(meltPoint = mean(meltPoint, na.rm = TRUE),
              plateau = mean(plateau, na.rm = TRUE),
              rSq = mean(rSq, na.rm = TRUE),
              n_rep = n()) %>%
    ungroup, 
  by = c("gene_name"), 
  suffix = c("_phospho", "_flowThrough"))

ggplot(phospho_combo_nn_df,
       aes(meltPoint_flowThrough, meltPoint_phospho)) +
  geom_point(alpha = 0.25, shape = 16) +
  geom_line(aes(x,y), data = tibble(x = 40:77.5, y = 40:77.5),
            color = "darkgray") +
  ggpmisc::stat_poly_eq(formula = y ~ x, parse = TRUE) +
  coord_cartesian(xlim = c(40, 77.5), ylim = c(40, 77.5)) +
  labs(x = expression('Tm[nbf] ('*~degree*C*')'), 
       y = expression('Tm[phospho] ('*~degree*C*')')) +
  theme_bw()
```


Now we join our results of the melting curve fitting for phosphopeptides and unmodified proteins and apply the test for differential melting points established by @savitski_2014
```{r, eval=FALSE}
nbf_tm_rep_fil <- out_nbf_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  group_by(gene_name) %>% 
  filter(n() > 2) %>% 
  ungroup %>% 
  dplyr::select(gene_name, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)

phospho_tm_rep_fil <- out_phospho_df %>% 
  filter(rSq > r2_thres, plateau < plateau_thres) %>% 
  group_by(gene_name, sequence, mod_sequence, phospho_site_STY) %>% 
  filter(n() > 2) %>% 
  ungroup() %>% 
  dplyr::select(gene_name, sequence, mod_sequence, phospho_site_STY, replicate, meltPoint) %>% 
  spread(replicate, meltPoint)

combo_tm_df <- left_join(phospho_tm_rep_fil, nbf_tm_rep_fil,
          by = "gene_name") %>% 
  rowwise() %>% 
  mutate(id = paste(gene_name, mod_sequence, phospho_site_STY, sep = "_")) %>% 
  ungroup()

combo_mean_tm_significant_df <- combo_tm_df %>%
  rowwise() %>%
  mutate(mean_phospho = mean(c(`1.x`, `2.x`, `3.x`, `4.x`, `5.x`), na.rm = TRUE),
         mean_nbf = mean(c(`1.y`, `2.y`, `3.y`, `4.y`, `5.y`), na.rm = TRUE)) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(delta_meltPoint_1 = `1.y` - `1.x`,
         delta_meltPoint_2 = `2.y` - `2.x`,
         delta_meltPoint_3 = `3.y` - `3.x`,
         delta_meltPoint_4 = `4.y` - `4.x`,
         delta_meltPoint_5 = `5.y` - `5.x`) %>%
  ungroup() %>%
  mutate(delta_meltPoint_1_scaled = scale(delta_meltPoint_1),
         delta_meltPoint_2_scaled = scale(delta_meltPoint_2),
         delta_meltPoint_3_scaled = scale(delta_meltPoint_3),
         delta_meltPoint_4_scaled = scale(delta_meltPoint_4),
         delta_meltPoint_5_scaled = scale(delta_meltPoint_5)) %>%
  mutate(p_value_1 = 1 - pnorm(abs(delta_meltPoint_1_scaled)),
         p_value_2 = 1 - pnorm(abs(delta_meltPoint_2_scaled)),
         p_value_3 = 1 - pnorm(abs(delta_meltPoint_3_scaled)),
         p_value_4 = 1 - pnorm(abs(delta_meltPoint_4_scaled)),
         p_value_5 = 1 - pnorm(abs(delta_meltPoint_5_scaled))) %>%
  ungroup() %>%
  mutate(p_adj_1 = p.adjust(p_value_1, method = "BH"),
         p_adj_2 = p.adjust(p_value_2, method = "BH"),
         p_adj_3 = p.adjust(p_value_3, method = "BH"),
         p_adj_4 = p.adjust(p_value_4, method = "BH"),
         p_adj_5 = p.adjust(p_value_5, method = "BH")) %>%
  rowwise() %>%
  mutate(significant = countAtLeast(x = c(p_adj_1, p_adj_2, p_adj_3, p_adj_4, p_adj_5),
                                    val_with_sign =
                                      c(delta_meltPoint_1_scaled,
                                        delta_meltPoint_2_scaled,
                                        delta_meltPoint_3_scaled,
                                        delta_meltPoint_4_scaled,
                                        delta_meltPoint_5_scaled),
                                    min_n =  2)) %>%
  ungroup()
```

To maximize overlap with @huang_2019 , we here choose gene names overlapping their data in case there are ambiguous gene names 
```{r}
## load Huang et al. phospho dataset
huang_phospho_tab <- readxl::read_xlsx(file.path("../dat", "41592_2019_499_MOESM4_ESM.xlsx"), skip = 2)

colnames(huang_phospho_tab)[15] <- "AveragePhosphoTm"
colnames(huang_phospho_tab)[21] <- "deltaTmPval"
colnames(huang_phospho_tab)[22] <- "HighQuality"

huang_phospho_hq <- filter(
  huang_phospho_tab,
  HighQuality == "Yes")

combo_mean_tm_fcs_choosen_gene_names <- 
  chooseCoherentGeneNames(in_df = combo_mean_tm_significant_df,
                          their_data = huang_phospho_tab)
  
```

Then we annotated the phosphorylation sites we found by mapping back the modified sequence to protein sequences obtained from Uniprot

```{r}
uniprot_seq_df <- read_csv("../dat/uniprot_seq_df.csv")

combo_mean_tm_pSite_anno <- left_join(
  combo_mean_tm_fcs_choosen_gene_names, 
  uniprot_seq_df,
  by = "gene_name") %>% 
  rowwise() %>% 
  mutate(Gene_pSite = getPhosphoSiteId(
    gene_name = gene_name, string = mod_sequence, seq = seq)) %>% 
  ungroup %>% 
  mutate(significant_huang = Gene_pSite %in% 
           filter(huang_phospho_hq, as.numeric(deltaTmPval) < 0.05)$Gene_pSite)

p_1b <- ggplot(combo_mean_tm_pSite_anno, aes(mean_nbf, mean_phospho)) +
  geom_point(shape = 16, alpha = 0.2, size = 0.5) +
  geom_line(aes(x,y), data = tibble(x = 40:65, y = 40:65),
            color = "darkgray", size = 0.25) +
  geom_point(data = filter(combo_mean_tm_pSite_anno, significant),
             color = "#b2182b",
             shape = 16, size = 0.5) +
  geom_point(data = filter(combo_mean_tm_pSite_anno, significant_huang),
             color = "#1f78b4",
             shape = 16, size = 0.5) +
  geom_point(data = filter(combo_mean_tm_pSite_anno, significant_huang, significant),
             color = "#ff7f00",
             shape = 16, size = 0.5) +
  coord_cartesian(xlim = c(40, 65), ylim = c(40, 65)) +
  ggpmisc::stat_poly_eq(formula = y ~ x, parse = TRUE, size = rsq_size) +
  labs(x = expression('T'[m]^unmodified* ' '* '('*~degree*C*')'), 
       y = expression('T'[m]^phospho* ' '* '('*~degree*C*')')) +
  theme_bw() +
  theme_paper
```

Now, we load the data on machine learning-derived functional scores for phosphosites by @ochoa_2020 and join them to our dataset

```{r}
# annotate ochoa et al. functional scores
ochoa_et_al_anno <- readxl::read_xlsx(file.path("../dat", "56035_2_data_set_478471_pxb8gk.xlsx"))
ochoa_et_al_mapping <- read_delim(file.path("../dat", "ochoa_mapping.txt"), delim = "\t")
ochoa_et_al_func_anno <- readxl::read_xlsx(file.path("../dat", "56035_2_data_set_478475_pxb8l1-tableS3.xlsx"))

ochoa_et_al_func_anno_psite <- left_join(
  ochoa_et_al_func_anno,
  ochoa_et_al_anno,
  by = c("uniprot", "position")
)

ochoa_et_al_anno_fil <- left_join(
  ochoa_et_al_func_anno_psite %>% dplyr::select(uniprot, position, residue, functional_score, is_disopred, isHotspot, isInterface),
  ochoa_et_al_mapping %>% dplyr::select(uniprot = From, gene_name = To),
  by = "uniprot") %>% 
  mutate(Gene_pSite = paste0(gene_name, "_", "p", residue, position))

combo_mean_tm_pSite_anno_functional_scores <- left_join(
  combo_mean_tm_pSite_anno %>% 
    rowwise() %>% mutate(mean_delta_meltPoint = mean(c(delta_meltPoint_1, delta_meltPoint_2, delta_meltPoint_3, 
                                                       delta_meltPoint_4, delta_meltPoint_5), na.rm = TRUE)) %>% 
    ungroup,
  ochoa_et_al_anno_fil %>% na.omit %>% dplyr::select(Gene_pSite, functional_score, is_disopred, isHotspot, isInterface),
  by = "Gene_pSite")

```

We join the functional scores also to the data by @huang_2019

```{r}
huang_raw_df <- readxl::read_xlsx(file.path("../dat", "41592_2019_499_MOESM3_ESM.xlsx"), skip = 2) 
colnames(huang_raw_df)[14] <- "AverageBulkTm"

huang_bulk <- huang_raw_df %>%
  dplyr::select(uniprot_id = `Uniprot Accession`,
                average_tm_nbf = AverageBulkTm) %>% 
  mutate(average_tm_nbf = as.numeric(average_tm_nbf))

huang_phospho_hq_potel_anno <- 
  huang_phospho_hq %>% 
  dplyr::select(uniprot_id = `Uniprot Accession`,
                Gene_pSite, 
                average_tm_phospho = AveragePhosphoTm,
                p_value = deltaTmPval) %>% 
  mutate(significant_potel = Gene_pSite %in% 
           filter(combo_mean_tm_pSite_anno_functional_scores, significant)$Gene_pSite) %>% 
  mutate(average_tm_phospho = as.numeric(average_tm_phospho),
         p_value = as.numeric(p_value)) %>% 
  left_join(huang_bulk, by = "uniprot_id")

p_1a <- ggplot(huang_phospho_hq_potel_anno, aes(average_tm_nbf, average_tm_phospho)) +
  geom_point(shape = 16, alpha = 0.2, size = 0.5) +
  geom_line(aes(x,y), data = tibble(x = 40:65, y = 40:65),
            color = "darkgray", size = 0.25) +
  geom_point(data = filter(huang_phospho_hq_potel_anno, p_value < 0.05),
             color = "#1f78b4", size = 0.5,
             shape = 16) +
  geom_point(data = filter(huang_phospho_hq_potel_anno, significant_potel),
             color = "#b2182b", size = 0.5,
             shape = 16) +
  geom_point(data = filter(huang_phospho_hq_potel_anno, p_value < 0.05, significant_potel),
             color = "#ff7f00", size = 0.5,
             shape = 16) +
  coord_cartesian(xlim = c(40, 65), ylim = c(40, 65)) +
  ggpmisc::stat_poly_eq(formula = y ~ x, parse = TRUE, size = rsq_size) +
  labs(x = expression('T'[m]^unmodified*' '* '('*~degree*C*')'), 
       y = expression('T'[m]^phospho*' '* '('*~degree*C*')')) +
  theme_bw() +
  theme_paper
```

And now we visualize the distributions of functional scores for the significantly vs. non-significantly differently thermally shifted phosphosites for the results by @huang_2019 and @potel_2020 
```{r}
# ochoa functional score on Huang et al data
huang_phospho_hq_functional_score <- left_join(
  huang_phospho_hq %>% 
    dplyr::select(Gene_pSite, average_deltaTm = AveragePhosphoTm,
                  p_value = deltaTmPval) %>% 
    mutate(average_deltaTm = as.numeric(average_deltaTm),
           p_value = as.numeric(p_value)) %>% 
    mutate(significant  =  (p_value < 0.05)),
  ochoa_et_al_anno_fil %>% na.omit %>% dplyr::select(Gene_pSite, functional_score),
  by = "Gene_pSite")

# combine both datasets to save space for plot
func_score_combo_df <- bind_rows(
  filter(combo_mean_tm_pSite_anno_functional_scores, !is.na(functional_score)) %>% 
    dplyr::select(Gene_pSite, significant, functional_score) %>% 
    mutate(group = "This study"),
  filter(huang_phospho_hq_functional_score, !is.na(functional_score)) %>% 
    dplyr::select(Gene_pSite, significant, functional_score)%>% 
    mutate(group = "Huang et al.")
)

p_1c <- ggplot(func_score_combo_df , aes(significant, functional_score)) +
  geom_violin(aes(fill = significant), alpha = 0.75, color = NA) +
  geom_boxplot(outlier.shape = NA, width = 0.25, alpha = 0.5) +
  ggsignif::geom_signif(comparisons = list(c("TRUE", "FALSE")),
                        textsize = anno_size, size = 0.25) +
  geom_text(stat = "count", aes(label = ..count..), y = 0.01,
            size = anno_size) +
  scale_fill_manual(values = c("gray10", "gray70")) +
  facet_wrap(~ group) +
  labs(x = "found significantly de/stabilized", y = "functional score by Ochoa et al.") +
  coord_cartesian(ylim = c(-0.1, 1.1)) +
  theme_bw() +
  theme(legend.position = "none", 
        strip.text = element_text(size = 8)) +
  theme_paper
```

We now plot examples melting curves showing significantly altered melting profiles in the phosphopeptide data compared to the unmodified proteins

```{r}
# examples from this study
p_1d <- plotDiffPhosphoMeltcurve(
  g_name = "LYN", p_seq = "_VIEDNEpYTAR_", 
  nbf_df = nbf_peptide_filtered,
  phospho_df = phospho_peptide_tab_filtered,
  Gene_pSite = "LYN_pY397") +
  theme_paper

p_1e <- plotDiffPhosphoMeltcurve(
  g_name = "LMNA", p_seq = "_SGAQASSpTPLpSPTR_", 
  nbf_df = nbf_peptide_filtered,
  phospho_df = phospho_peptide_tab_filtered,
  Gene_pSite = "LMNA_pT19pS22") +
  theme_paper

p_1f <- plotDiffPhosphoMeltcurve(
  g_name = "CARHSP1", p_seq = "_GNVVPpSPLPTRR_", 
  nbf_df = nbf_peptide_filtered,
  phospho_df = phospho_peptide_tab_filtered,
  Gene_pSite = "CARHSP1_pS41") +
  theme_paper

```

Finally, we combine all plots created above and plot them together
```{r}
# assemble full figure
cowplot::plot_grid(
  cowplot::plot_grid(p_1a, p_1b, p_1c, 
         labels = c("a", "b", "c"), ncol = 3, rel_widths = c(0.3, 0.3, 0.3)),
  cowplot::plot_grid(p_1d, p_1e, p_1f, labels = c("d", "e", "f"), 
            ncol = 3, rel_widths = c(0.3, 0.3, 0.3)),
  ncol = 1
)

```

## Make output tables
```{r eval=FALSE}
supp1_tab <- phospho_peptide_tab_filtered %>% 
    dplyr::select(gene_name, protein_id, 
                  sequence, modifications,
                  key, replicate, temperature) %>% 
    unite(combo_var, c(key, replicate, temperature), 
          sep = "_")

write_csv(supp1_tab)

supp2_tab <- nbf_peptide_filtered  %>% 
    ungroup %>% 
    dplyr::select(gene_name, protein_id, 
                  sequence, modifications,
                  score,
                  key, replicate, temperature,
                  value) %>% 
    unite(combo_var, c(key, replicate, temperature), 
          sep = "_") %>% 
    spread(combo_var, value)

write_csv(supp2_tab)
```


## Additonal comparisons of the two datasets

```{r}
delta_tm <- left_join(
    combo_mean_tm_pSite_anno %>% 
        dplyr::select(Gene_pSite, mean_delta_meltPoint),
    huang_phospho_hq_potel_anno %>% 
        mutate(huang_phospho_hq_potel_anno = average_tm_phospho - 
                   average_tm_nbf) %>% 
        dplyr::select(Gene_pSite, mean_delta_meltPoint_huang, p_value),
    by = "Gene_pSite"
) %>% 
    na.omit()

ggplot(delta_tm, aes(mean_delta_meltPoint_huang, mean_delta_meltPoint)) +
    geom_point(alpha = 0.5) +
    geom_point(color = "red", data = filter(delta_tm, p_value < 0.05)) +
    geom_smooth(method = "lm") +
    # ggrepel::geom_text_repel(
    #     aes(label = Gene_pSite),
    #     data = filter(delta_tm, p_value < 0.05,
    #                   Gene_pSite %in% c("ANLN_pS54", "TPI1_pS58",
    #                                     "RPRD2_pS1099", "BAG6_pS964pS973"))) +
    labs(x = expression(Delta*' T'[m]^Potel*' '* '('*~degree*C*')'), 
         y = expression(Delta* ' T'[m]^Huang*' '* '('*~degree*C*')')) +
    theme_bw()
```

# References